CH 7: Interactive Matrix Language (ML)

R is a Matrix language, and it is important to know how to use matrices and 
translate our R understanding into SAS programming. This chapter should focus
on highlighting the basics of Matrix understanding, how to work with Modules and 
subroutines, calling SAS Data Sets and Procedures and finally simulating results

in R, we use the following code to create a matrix and conduct operations. 

```{r}
# Create Matrix 
x = matrix(c(1,2,3,4),nrow = 2,ncol = 2,byrow = T)
y = matrix(c(5,6,7,8),nrow = 2,ncol = 2,byrow = T)

# Element wise Operations
x ^ 2
x * y 

# Matrix Operations 
x %% y 

# Accessing Elements 
x[1,]
y[1,1:2]

# Reduction 
max(x)
sum(x)
colMeans(x)


```


PROC IML;
  IML syntax
QUIT; 


in SAS:
{} Making a Matrix from Literal Values; 
[] Referring to Matrix Subscripts, specifying options; 
() Controlling Order of Operations, Functions 

Creating a Module: 

```{r}

# function.name = function(arg1, arg2,...){
#  statements
#  return(arguments)
# }
#
#
#

```

in SAS, the syntax is as follows: 

START name <(argument1, argument2,...)>
  <GLOBAL(arguement1, arguement2,...)>;
  statements;
  <RETURN(matrix);>
FINISH; 


PROC IML;
	start standarize(x);
		n=nrow(x);
		mean=x[:,];
		xbar=repeat(mean,n,1);
		x=x-xbar;
		stdv=std(x);
		x=x/stdv;
		return(x);
	finish;
	
	n = 10;
	call randseed(800);
	mymat = randfun(n, "Normal", 5,5)
		 || randfun(n,"Uniform", 10,15)
		 || randfun(n, "Exponential", 7);
	print mymat;
	stand = standardize(mymat); 
	print(stand);
quit; 

PROC IML;
	start standsub(stand, mean, stdv, x);	
		n=nrow(x);
		mean=x[:,];
		xbar=repeat(mean,n,1);
		x=x-xbar;
		stdv=std(x);
		stand=x/stdv;
	finish;
	
	n = 10; 
	call randseed(800);
	mymat = randfun(n, "Normal", 5,5)
		 || randfun(n,"Uniform", 10,15)
		 || randfun(n, "Exponential", 7);
	call standsub(standardized,m,s,mymat);
	print m,s,standardized;
quit; 


Simulations: 

1.) Obtain Solutions 
2.) Evaluate Methods

```{r}
# Sample Distribution 
SimpleReg = function(n){
  
  beta0 = 5
  beta1 = 2
  xvals = runif(n)*n
  error = rnorm(n, mean = 0, sd = 5)
  y = beta0 + beta1*xvals + error
  
  x = cbind(rep(1,n), xvals)
  beta = solve(t(x) %*% x, t(x) %*% y)
  return(beta[2])
}

#simulate
reps = 1000000
samp = rep(0, reps)
for(i in 1:reps){
  samp[i] = SimpleReg(20)
}

#Analyze
hist(samp,50)
mean(samp)
sd(samp)

```




Monty Hall Simluation in SAS; 

/*SP4R07d06*/

/*Part A*/
proc iml;
   numberIterations=10000;
   call randseed(802);

/*Part B*/
   do iteration=1 to numberIterations;
      doors = {1 2 3};
      carDoor=sample(doors,1);

/*Part C*/
     *Pick door for Monty Hall to open;
     if carDoor=1 then openDoor=randfun(1,"Bernoulli",.5) + 2;
     else if carDoor=2 then openDoor=3;
     else if carDoor=3 then openDoor=2;

/*Part D*/
     *Determine door for switching strategy;
     if openDoor=2 then switchDoor=3;
     else if openDoor=3 then switchDoor=2;

/*Part E*/
     *Determine which strategy wins;
     if carDoor=1 then stayWin=1;
     else stayWin=0;

     if carDoor=switchDoor then switchWin=1;
     else switchWin=0;
	 /*switchWin=carDoor=switchDoor;*/

/*Part F*/
     *Collect results to a single matrix;
     results=results // (iteration || carDoor || openDoor || stayWin || switchWin);
   end;

/*Part H*/
   reset noname;
   resultsSubset = results[1:10,];
   print resultsSubset [colname={iteration carDoor openDoor 
      stayWin switchWin}];

   percentageWins=results[:,{4 5}];
   print percentageWins [colname={stay switch}];
quit;










