SAS vs R in Data Types 

R makes it very simple to read in data, while R stores all of the environmental
and global data into memory, SAS stores the data in storage, and then calls
when needed. 

The following SAS code accomplishes viewing example data and creating a SAS df. 

/*SP4R02d01*/

/*Part A*/
data sp4r.example_data;
   length First_Name $ 25 Last_Name $ 25;
   input First_Name $ Last_Name $ age height;
   datalines;
   Jordan Bakerman 27 68
   Bruce Wayne 35 70
   Walter White 51 70
   Henry Hill 65 66
   JeanClaude VanDamme 55 69
;
run;

/*Part B*/
data sp4r.example_data2;
   length First_Name $ 25 Last_Name $ 25;
   input First_Name $ Last_Name $ age height @@;
   datalines;
   Jordan Bakerman 27 68 Bruce Wayne 35 70 Walter White 51 70
   Henry Hill 65 66 JeanClaude VanDamme 55 69
   ;
run;


The same task in R is as follows: 

```{r}
#Data in Lists 
First_Name <- list("Jordan", "Bruce", "Walter", "Henry", "Jean Claude")
Last_Name <- list("Bakerman", "Wayne", "White", "Hill", "VanDamme")
Age <- list(27, 35, 51, 65, 55)
Height <- list(68, 70, 70, 66, 69)

# Convert Lists into DF columns
df <- data.frame(unlist(First_Name), unlist(Last_Name), unlist(Age), unlist(Height))

names(df) <- c("First", "Last", "Age", "Height")

cat("The Dataframe is :\n")
print(df)

```


The next code chunk deals with reading in Data, Formatting. 

If I wanted to import Data into SAS, I need to use a DATA step, as follows: 

*** PAY SPECIAL ATTENTION TO THE "INFILE" syntax *** 

/*SP4R02d02*/

/*Import data using a DATA step*/
data sp4r.all_names;
   length First_Name $ 25 Last_Name $ 25;
   infile "&path/allnames.csv" dlm=',';
   input First_Name $ Last_Name $ age height;
run;

/***************************************************************************************************/
/*Import data using PROC IMPORT*/
proc import out=sp4r.baseball 
   datafile= "&path/baseball.csv" DBMS=CSV REPLACE;
   getnames=yes;
   datarow=2; 
run;

/*Rename the variables*/
data sp4r.baseball;
   set sp4r.baseball;
   rename nAtBat = At_Bats
	     nHits = Hits
	     nHome = Home_Runs
	     nRuns = Runs
	     nRBI = RBIs
	     nError = Errors;
run;

/***************************************************************************************************/
/*Creating a SAS data set from delimited data by hand*/
data sp4r.example_data3;
   length First_name $ 25;
   infile datalines dlm='*';
   input First_Name $ Last_Name $ age height;
   datalines;
   Jordan*Bakerman*27*68
   Bruce*Wayne*35*70
   Walter*White*51*70
   Henry*Hill*65*66
   Jean Claude*Van Damme*55*69
;
run;

Similar code in R is fairly easy to figure out
```{r}

```

formatting in SAS is where things can get tricky, 

/*SP4R02d03*/

data employees;
   input name $ bday :mmddyy8. @@;
   datalines;
   Jill 01011960 Jack 05111988 Joe 08221975
   ;
run;

proc print data=employees;
run;

data employees;
   input name $ bday :mmddyy8. @@;
   format bday mmddyy10.;
   label name="First Name" bday="Birthday";
   datalines;
   Jill 01011960 Jack 05111988 Joe 08221975
   ;
run;

proc print data=employees label;
run;


similar code in R: 

```{r}

```












