Chapter 5: Descriptive Procedures, Output Delivery System, and Macros

The level of difficulty increases dramatically here. 

While R stores objects in memory, SAS does not. We need to figure out a way 
to create global variables that can be passed into any procedure or DATA step. 

Creating macro programs in R (essentially a functon) which will increase the 
customization of SAS code. These macros will generate and execute DATA and PROC
steps automatically based on parameters we pass in the macro call. 

In R, we use code similar to this to find correlation and co-variance 

```{r}
require(stats); require(graphics)
# Compute the correlation and variance matrices: 

cor(cbind(cars$Horsepower, cars$weight, cars$length))
cov(cbind(cars$Horsepower, cars$weight, cars$length))

# Create freqency tables
table(cars$Origin)
table(cars$Type)

# Create contigency tables
table(cars$Origin, cars$Type)

#Create QQ-plot
qqnorm(cars$Hits)

# Summary Statistics: 

mean(cars$mpg_city)
median(cars$mpg_city)
mode(cars$mpg_city)
range(cars$mpg_city)
var(cars$mpg_city)
sd(cars$mpg_city)
sum(cars$mpg_city)
min(cars$mpg_city)
max(cars$mpg_city)
quantile(cars$mpg_city, c(0.1, 0.5, 0.1, 0.25, 0.5, 0.75, 0.05, 0.9, 0.99))
# Alternatively just use this: 

summary(cars$mpg_city)

```


for corr table: 

proc corr data = sp4r.cars corr; 
  var horsepower weight length; 
run; 

All other summary level / relational schema will run similarly. 

Below, we see SAS code for running frequency tables on a certian variable, 
as well as getting summary statistics, advanced statistical outcomes, and 
overlaying a plot with normalized density. 

/*SP4R05s01*/

/*Part A*/
proc freq data=sp4r.ameshousing;
   tables central_air house_style / plots=freqplot;
run;

/*Part B*/
ods select pearsoncorr;
proc corr data=sp4r.ameshousing;
   var saleprice garage_area basement_area gr_liv_area;
run;

/*Part C*/
ods output summary=summary_table;
proc means data=sp4r.ameshousing p10 median p90;
   var saleprice gr_liv_area;
   class yr_sold;
run;

proc print data=summary_table;
run;

/*Part D*/
proc univariate data=sp4r.ameshousing;
   var gr_liv_area;
   histogram gr_liv_area / normal kernel;
   qqplot gr_liv_area / normal(mu=est sigma=est);
   output out=gr_percs pctlpts= 40 to 60 by 2 pctlpre=gr_;
run;

proc print data=gr_percs;
run;


* Selecting OUtput to print and supress * 


in R, we use objects to create models and define outputs. 


```{r}
mylm = lm(y~x)

# print default output
print(mylm)


# Finding what is in the object
names(mylm)

# Pull desired outputs from the object
mylm$residuals

# create dataframe from output

mylm_res = data.frame(mylm$residuals)


```



in SAS, we use the ODS (output delivery system).
We can get ~inside~ the ODS to modify the appearance of an output, changing the
output destination, selecting specific outputs, and create new tables 

/*SP4R05d03*/

/*Part A*/
ods select basicmeasures;
ods output basicmeasures = sp4r.SalePrice_BasicMeasures;
proc univariate data=sp4r.ameshousing;
   var saleprice;
run;

proc print data=sp4r.saleprice_basicmeasures;
run;

/*Part B*/
proc univariate data=sp4r.ameshousing;
   var saleprice;
   output out=sp4r.stats mean=saleprice_mean pctlpts= 40, 45, 50, 55, 60 
      pctlpre=saleprice_;
run;

proc print data=sp4r.stats;
run;

/*Part C*/
proc means data=sp4r.ameshousing;
   var saleprice garage_area;
   output out=sp4r.stats mean(saleprice)=sp_mean median(garage_area)=ga_med;
run;

proc print data=sp4r.stats;
run;

/*Part D*/
proc means data=sp4r.ameshousing;
   var saleprice garage_area;
   output out=sp4r.stats mean= std= / autoname;
run;

proc print data=sp4r.stats;
run;


MACROS: 
All Macro variables are by default Global. Just like creating a personal function 
in R, 

First tip with assigning Macros is to use the "&"  

To use PROC SQL to help automate processes: 

/*SP4R05d04*/

/*Part A*/
proc means data=sp4r.ameshousing;
   var saleprice;
   output out=sp4r.stats mean=sp_mean std=sp_sd;
run;

proc sql;
   select sp_mean into :sp_mean from sp4r.stats;
   select sp_sd into :sp_sd from sp4r.stats;
quit;

/*Part B*/
data sp4r.ameshousing;
   set sp4r.ameshousing;
   sp_stan = (saleprice - &sp_mean) / &sp_sd;
run;

proc print data=sp4r.ameshousing (obs=6);
   var saleprice sp_stan;
run;

proc means data=sp4r.ameshousing mean std;
   var saleprice sp_stan;
run;

/*Part C*/
proc contents data=sp4r.cars varnum out=carscontents;
run;

proc print data=carscontents;
   var name type;
run;

/*Part D*/
proc sql;
   select distinct name into: vars_cont separated by ' ' from carscontents where type=1;
   select distinct name into: vars_cat  separated by ' ' from carscontents where type=2;
quit;

%put The continuous variables are &vars_cont and the categorical variables are &vars_cat;


The & Put statement helps store data, and will print as is. We do not need to include 
this in the DATA or Proc step. 

Creating MACRO Programs (Running Repetitive SAS programs)


For instance if I wanted to generate a random number of observations, and then 
use that information to run and generate a report, I can use R code that is 
similar to the following: 

```{r}
randnorm = function(n){
   
   vec = rnorm(n)
   # Summary statistics 
   mean(vec)
   median(vec)
   sd(vec)
   min(vec)
   max(vec)
   
   # summary(vec) # Alternative to the above lines 
   
   # Create Plots: 
   par(mfrow = c(1,2))
   hist(vec)
   plot(vec, type = "b")
}

# Call Function passing in Parameter N
randnorm(n = 10000)

```

With Macros, Our goals are as follows: 
1.) Be able to use symbolic substitution within SAS code (&)
2.) Automate production of SAS code 
3.) Write If / Then / Else conditional SAS code 


Creating a Macro: 

% Macro today; 
   % put Today is &sysday &sysdate9;
% mend; 

%today 









